---
title: "Lab 4"
author: "Stephanie Gluck"
date: "2/10/2020"
output: 
  html_document: 
    theme: cerulean
---

```{r setup}

library(rio)
library(tidyverse)
library(ggplot2)
library(here)
library(colorblindr)
library(scales)

#saving these for future reference 
#devtools::install_github("wilkelab/cowplot")
#install.packages("colorspace", repos = "http://R-Forge.R-project.org")
#devtools::install_github("clauswilke/colorblindr")
#remotes::install_github("tidyverse/ggplot2")
#install.packages("colorblindr")

theme_set(theme_minimal())
```

```{r load_data}

pth <- here("honeyproduction.csv")
d <- import(pth)

state <- rio::import("https://github.com/cphalpert/census-regions/raw/master/us%20census%20bureau%20regions%20and%20divisions.csv") 

```

## Honey Visualization 

Total produciton of honey (million pounds) across years by state. Using color to highlight the west coast (Washington, Oregon, and California).

```{r plot1, fig.width = 8, fig.height = 7}
#head(d)

# this is not really needed unless you just want it for better viewing when
# printing your data frame to the console.
d <- arrange(d, -totalprod) 

p <- ggplot(d, aes(fct_rev(as_factor(state)), totalprod/10^6)) + 
  geom_col(fill = "#a6611a") + 
  geom_col(data = filter(d, 
                         state == "OR" |
                           state == "WA"|
                           state == "CA"), 
           fill = "#018571") +
  scale_y_continuous(labels = scales::unit_format(unit = "M"), 
                     expand = c(0, 0, 0.05, 0)) +
  theme(panel.grid.major.y = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.title.position = "plot") +
  coord_flip() + 
  labs(x = "",
       y = "Honey Production (million pounds)",
       title = "Total Honey Production by State")
p
# This looks really nice!
```

Reproduce the plot according three different kinds of color blindness, as well as a desaturated version.

```{r plot1_colorblind_test}

colorblindr::cvd_grid(p)

```

Color blind safe palette 

```{r plot1_colorblind_safe, fig.width = 8, fig.height = 7}

p1_colorsafe <- ggplot(d, aes(fct_rev(as_factor(state)), totalprod/10^6)) + 
  geom_col(fill = "#1f78b4") + 
  geom_col(data = filter(d, 
                         state == "OR" |
                           state == "WA"|
                           state == "CA"), 
           fill = "#a6cee3") +
  theme(panel.grid.major.y = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.title.position = "plot") +
  scale_y_continuous(labels = unit_format(unit = "M"), 
                     expand = c(0, 0, 0.05, 0)) +
  coord_flip() + 
  labs(x = "",
       y = "Honey Production (million pounds)",
       title = "Total Honey Production by State")

p1_colorsafe # Ooh.. I like this one even better

```

Check plot for colorblind friendliness 

```{r plot1_colorblind_safe_test}

colorblindr::cvd_grid(p1_colorsafe)
# pretty darn good! Even in the desaturated version.
```


## Honey State Visualization 

Join file 

```{r join_file}

state <- state %>% rename(state = `State Code`)

d_state <- left_join(d, state)

#head(d_state)

```


Total produciton of honey (million pounds) across years by state. Using color to highlight region of the country the state is from. 

```{r plot2, fig.width = 8, fig.height = 7}

ggplot(d_state, aes(reorder(state, totalprod), totalprod/10^6)) + 
  geom_col(aes(fill = Region)) +
  scale_fill_OkabeIto() +
  scale_y_continuous(labels = unit_format(unit = "M"), 
                     expand = c(0, 0, 0.05, 0)) +
  theme(panel.grid.major.y = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        plot.title.position = "plot") +
  coord_flip() + 
  labs(x = "",
       y = "Honey Production (million pounds)",
       title = "Total Honey Production by State") 

```

**Summary**: A bulk majority of the country's honey is produced by North Dakota, California and South Dakota and two of these states are in the Midwest region (ND & SD) of the United States. Honey production for ND and SD combined is around 700 million pounds with CA produciton at about 350 million pounds. The region with the smallest total honey production is the Northeast of the United States. 

<!-- Perfect summary! -->

## Honey Production Heatmap by Region 

Average honey production across years by region

```{r heatmap, fig.width = 8, fig.height = 6}

d_heat <- d_state %>% 
  group_by(Region, year) %>%  # Each new function should go on a new line
  summarise(production = mean(totalprod))

#head(d_heat)

ggplot(d_heat, aes(year, Region)) + 
  geom_tile(aes(fill = production/10^6)) + 
  theme_minimal(base_size = 15) +
  scale_fill_viridis_c(option = "magma") +
  coord_fixed() + # I think this helps
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5), 
        legend.title.align = 0.5) +
  labs(x = "",
       y = "",
       title = "Average Honey Production Across Years by Region", 
       fill = "Honey Production \n(million pounds)")

```

## Additional Honey Production Visualization 

```{r plot3, fig.width = 8, fig.height = 6}

ggplot(d_heat, aes(year,production/10^6, color = Region)) + 
  geom_line(size = 1.5) + 
  geom_point(size = 2.5) +
  theme_minimal(base_size = 15) +
  scale_color_viridis_d(name = "") + 
  theme(panel.grid.major.y = element_blank(), 
        plot.title.position = "plot") +
  scale_y_continuous(labels = unit_format(unit = "M", accuracy = 1), 
                     expand = c(0.05, 0, 0.05, 0)) +
  labs(x = "",
       y = "Honey Production (million pounds)",
       title = "Average Honey Production Across Years by Region") + 
  geom_text(data = filter(d_heat, year == "2012"), 
            aes(y = production/10^6, label = Region), 
            color = "gray20", 
            nudge_x = 0.3, 
            hjust = 0, 
            size = 5) + 
  guides(color = "none") + 
  xlim(NA, 2014)
# Very nice!
```

